---
title: "Finding marginal seats in the 2017 election"
runtime: shiny
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    social: menu
    source_code: https://github.com/nickforr/ukEelctions
    theme: simplex

---

```{r global, include=FALSE}
library(flexdashboard)
library(shiny)
library(raster) #Need to put raster before tidyverse to avoid masking select
library(tidyverse)
library(readxl)
library(scales)
library(hrbrthemes)
library(viridis)
library(leaflet)
library(waffle)

if (!file.exists("electionResults2015.xlsx")) {
  download.file("http://www.electoralcommission.org.uk/__data/assets/excel_doc/0011/189623/2015-UK-general-election-data-results-WEB.xlsx", "electionResults2015.xlsx")
}

simplifiedConstituencyShapefile <- 
  readRDS("simplifiedShapefile.RDS")

if (!file.exists("votingData.RDS")) {
  candidates <- 
    readxl::read_excel("electionResults2015.xlsx", sheet = 2) %>%
    filter(PANO != is.na(PANO)) %>%
    rename(constituencyId = `Constituency ID`)
  constituency <- 
    readxl::read_excel("electionResults2015.xlsx", sheet = 3) %>%
    filter(PANO != is.na(PANO)) %>%
    rename(
      constituencyId = `Constituency ID`,
      constituencyName = `Constituency Name`
    )
  votingData <- 
    left_join(
      candidates %>%
        select(PANO, Surname, constituencyId, 
          `Party abbreviation`, Votes),
      constituency %>%
        select(PANO, constituencyId, constituencyName, 
          Electorate, `Valid Votes`, Region),
      by = c("PANO", "constituencyId")
    ) %>% 
    filter(PANO != is.na(PANO)) %>%
    rename(
      validVotes = `Valid Votes`,
      partyAbbreviation = `Party abbreviation`
    ) %>%
    group_by(PANO) %>%
    mutate(resultRank = row_number(desc(Votes))) %>%
    filter(resultRank <= 4) %>%
    mutate(
      votesPpnElectorate = Votes / Electorate,
      VotesPpnVoted = Votes / validVotes,
      votesPpnLabel = sprintf("%.0f%%", VotesPpnVoted * 100),
      winningVotes = max(Votes),
      shortfallVotes = winningVotes - Votes,
      shortfallVotesPpnElectorate = shortfallVotes / Electorate,
      shortfallVotesPpnVoted = shortfallVotes / validVotes
    ) %>%
    ungroup() %>%
    mutate(
      partyAbbreviation = 
        factor(partyAbbreviation, levels = unique(partyAbbreviation)))
  
  votingData <- 
    votingData %>%
    left_join(
      votingData %>%
        filter(resultRank == 2) %>%
        mutate(shortfallRank = row_number(shortfallVotesPpnVoted)) %>%
        group_by(Region) %>%
        mutate(shortfallByRegionRank = row_number(shortfallVotesPpnVoted)) %>%
        ungroup() %>%
        select(PANO, shortfallRank, shortfallByRegionRank)
    )
  rm(candidates, constituency)
  saveRDS(votingData, "votingData.RDS")
} else {
  votingData <- readRDS("votingData.RDS")
}

allPartyNames <- 
  unique(filter(votingData, resultRank == 1)$partyAbbreviation)
 
partyPalette <- 
  viridis(length(allPartyNames), direction = -1) %>%
  set_names(allPartyNames)

partySelection <- 
  c("Con", "Lab", "UKIP", "SNP", "LD", "Green", "PC", "Ind") 
selectionPalette <- 
  c("#0087dc", "#d50000", "#B3009D", "#FFF95D", "#FDBB30", "#8dc63f", 
    "#3F8428", "#d18975")
partyPalette[partySelection] <- selectionPalette
    
mapVotingData <- 
  votingData %>%
  filter(resultRank <= 3) %>% 
  mutate(mapLabel = 
      sprintf("%s (%.0f%%)", as.character(partyAbbreviation), 
        VotesPpnVoted * 100)) %>%
  select(constituencyId, constituencyName, resultRank, mapLabel, 
    shortfallRank) %>%
  spread(key = resultRank, value = mapLabel) %>%
  left_join(
    votingData %>%
      filter(resultRank == 2) %>%
      select(constituencyName, shortfallVotesPpnVoted),
    by = c("constituencyName")
  )
simplifiedConstituencyShapefile@data <- 
  left_join(
    simplifiedConstituencyShapefile@data, 
    mapVotingData, 
    by = c("CODE" = "constituencyId")
  )
mapLabels <- 
  sprintf(
    "<strong>%s</strong><br>
    First: %s<br>
    Second: %s<br>
    Third: %s",
    simplifiedConstituencyShapefile$constituencyName,
    simplifiedConstituencyShapefile$`1`, 
    simplifiedConstituencyShapefile$`2`,
    simplifiedConstituencyShapefile$`3`
  ) %>% lapply(htmltools::HTML)

regionList <- sort(unique(votingData$Region))

leafletPalette <- 
    colorNumeric("viridis", 
      domain = simplifiedConstituencyShapefile$shortfallVotesPpnVoted,
      reverse = TRUE
    )

waffleData2015 <- 
  votingData %>%
  filter(resultRank == 1) %>%
  group_by(partyAbbreviation) %>%
  summarise(a = n()) %>%
  arrange(desc(a))

waffle2015 <- waffle::waffle(setNames(waffleData2015$a, waffleData2015$partyAbbreviation), colors = partyPalette, rows = 14, 
  equal = TRUE, size = 1, legend_pos = "bottom") 
```

Sidebar {.sidebar}
=======================================================================

Whether you have views on Scottish independence or on Brexit, the upcoming election suggests a greater relevance for tactical voting.  

Use the [Explore constituency](#section-explore-constituency) tab to see more detail about your chosen constituency. The [Marginal constituencies in region](#section-marginal-constituencies-in-region) tab highlights the top marginal contituencies in the selected region.  And the [Overall position](#section-overall-position) tab considers the combined picture.

```{r}

#Select input for region
selectInput(
  inputId = "regionChoice", label = "Select a region:", 
  choices = c("All", regionList), selected = "Scotland"
)

#Filter data based on region data
regionVotingData <- reactive({
  if (input$regionChoice == "All") {
    votingData
  } else {
    votingData %>%
      filter(Region == input$regionChoice)
  }
})

constituenciesInRegion <- reactive({
  regionData <- regionVotingData()
  sort(unique(regionData$constituencyName))
})

#render ui to select constituecy
output$constituencySelect <- renderUI({
  constituencies <- constituenciesInRegion()
  selectInput(
    inputId = "constituencyChoice", label = "Select constituency:",
    choices = constituencies
  )
})
uiOutput("constituencySelect")

constituencyVotingData <- reactive({
  req(input$constituencyChoice)
  votingData %>% 
    filter(constituencyName == input$constituencyChoice)
})
```

<hr>
Polling suggests that there may be a *few* changes from the last election results, hence some simple'what if' analysis.  
The slider below lets you explore what might happen were the Labour vote share to fall by the specified amount:
```{r}
sliderInput(
  inputId = "laboutLeavers",
  label = NULL,
  min = 0, max = 75, value = 25, step = 5,
  post = "%"
)
```
I have assumed that half of those deserting Labour won't vote at all, whilst the other half are split evenly across Conservative and Liberal Democrat.  
In addition, the checkkbox below allows you to look at what would happen if all UKIP voters defect with 80% voting Conservative and 20% voting Labour. 
```{r}
checkboxInput(
  inputId = "ukipWipedOut", 
  label = "UKIP wiped out?",
  value = TRUE
)
```
Clearly my assumptions are gross over-simplifications but the hope is that they might be interesting...

Explore constituency
=======================================================================

Column
-----------------------------------------------------------------------

### Constituency results

```{r}
output$constituencyChoice <- renderText({
  input$constituencyChoice
})
output$constituencyPlot <- renderPlot({
  constituencyChoice <- input$constituencyChoice
  constituencyData <- constituencyVotingData()
  
  constituencyData %>%
    ggplot() +
    theme_minimal(base_size = 16) +
    labs(
      subtitle = constituencyChoice, 
      caption = 
        paste0(
          scales::ordinal(unique(constituencyData$shortfallByRegionRank)), 
          " most marginal seat in region (",
          scales::ordinal(unique(constituencyData$shortfallRank)), 
          " overall)"
        )) +
    geom_col(
      aes(
        x = partyAbbreviation, 
        y = Votes,
        fill = partyAbbreviation
      )
    ) +
    geom_text(
      aes(
        x = partyAbbreviation, y = Votes, label = votesPpnLabel
      ), size = 4, vjust = 0
    ) +
    scale_fill_manual(values = partyPalette, guide = "none") +
    scale_y_continuous(
      labels = comma, breaks = pretty_breaks(n = 6), expand = c(0.1, 0)) +
    xlab(NULL) + ylab(NULL) + 
    theme(plot.caption = element_text(size = 12))
})
plotOutput("constituencyPlot")
```

### 'What if' constituency results

Something here

Column
-----------------------------------------------------------------------

### Mapping constituencies

```{r}
output$constituencyMap <- renderLeaflet({
  #req(input$regionChoice)

  leaflet(data = simplifiedConstituencyShapefile) %>% 
    addProviderTiles(providers$CartoDB.Positron) %>%
    # setView(
    #   mean(selectedData@bbox[1, ]), mean(selectedData@bbox[2, ]), 
    #   7) %>%
    addPolygons(
      fillColor = ~leafletPalette(shortfallVotesPpnVoted), 
      fillOpacity = 0.6, 
      layerId = ~POLYGON_ID,
      stroke = TRUE, color = "white", 
      dashArray = 3, 
      weight = 1, 
      label = mapLabels,
      labelOptions = labelOptions(
        style = list("font-weight" = "normal", padding = "3px 5px"),
        textsize = "10px", direction = "auto")
    ) %>%
    addLegend(
      pal = leafletPalette, values = ~shortfallVotesPpnVoted, 
      labFormat = labelFormat(suffix = "%", transform = function(x) 100 * x),
      opacity = 0.7, title = "Margin of <br>victory", position = "bottomright")
})
# Click event for the map (will use to generate chart)
# click_constituency <- eventReactive(input$constituencyMap_shape_click, {
#     x <- input$constituencyMap_shape_click
#     x$id
#   })
observe({
  req(input$constituencyChoice) # do this if click_tract() is not null
  # Add the clicked tract to the map in black, and remove when a new one is
  # clicked
  leafletProxy('constituencyMap') %>%
    removeShape('selectedConstituency') %>%
    addPolygons(
      data = simplifiedConstituencyShapefile[simplifiedConstituencyShapefile$constituencyName == input$constituencyChoice, ], 
      fill = FALSE, color = "black", weight = 3,
      opacity = 1, layerId = 'selectedConstituency')
  })

leafletOutput("constituencyMap")
```


Marginal constituencies in region
=======================================================================

### Top marginal constituencies in selected region

```{r}
output$marginalFacet <- renderPlot({
  regionData <- regionVotingData()
  regionData %>%
    filter(shortfallByRegionRank <= 6) %>%
    ggplot() +
    theme_minimal(base_size = 16) +
    labs(
      caption = 
        paste0("Showing consitituencies in ", unique(regionData$Region))
    ) +
    geom_col(
      aes(
        x = partyAbbreviation, 
        y = Votes,
        fill = partyAbbreviation
      )
    ) +
    geom_text(
      aes(
        x = partyAbbreviation, y = Votes, 
        label = sprintf("%.0f%%" , VotesPpnVoted * 100)
      ), size = 4, vjust = 0
    ) +
    scale_fill_manual(values = partyPalette, guide = "none") +
    scale_y_continuous(
      labels = comma, breaks = pretty_breaks(n = 6), expand = c(0.1, 0)) +
    xlab(NULL) + ylab(NULL) + 
    theme(plot.caption = element_text(size = 12)) +
    facet_wrap(~constituencyName, ncol = 3)
})
plotOutput("marginalFacet")
```

Overall position {data-orientation=columns}
=======================================================================

Column
-----------------------------------------------------------------------
### 2015 results

```{r}
iron(waffle2015, waffle2015)
```


Column
----------------------------------------------------------------------

### What if

```{r}
knitr::kable(waffleData2015)
```


About
=======================================================================

### Background  

I put together this app partly because I like playing around with [R](https://cran.r-project.org) and [Shiny](http://shiny.rstudio.com) and also because the upcoming 2017 UK election is likely to be extremely disappointing, unless you're in favour of an extreme Brexit.  

As someone who sees themselves as Scottish, British and European, I have little time for the parochialism and, in many cases, racism on show in both the Scottish Independence and 'Brexit' debates; and even less time for the hypocrisy (and, frankly, dishonesty) shown by many politicians regarding any uncomfortable information that might impact negatively on their views.

When it comes to Scotland, it's my view that independence would do far greater harm to the poorest in our society than any austerity measures that the Conservative party in Westminster may enact; there are degrees of harm and, as unfortunate as it is, voting to cut off your foot is better than voting to cut off your leg. With regard to Europe, I think the blame for how many people feel towards the EU should sit with our own politicians long before we start looking at others. 

If pushed to put a label on it, I would probably characterise my politics as centre-left; I'm often torn between the Labour party and the Liberal Democrats. But whilst I may disagree with many of their policies, I would note that there are many Conservative MPs whom I respect (the attempt to paint the Conservative party as 'evil monsters' by many in Scotland does nobody any good).

More generally, I would prefer to see far less tribalism and a much greater willingness to seek compromise and collaboration across party lines.

### References

Data sourced from: http://www.electoralcommission.org.uk

Constituency maps contain OS data &copy; Crown copyright and database right 2017

Flexdashboard layout inspiration: <https://github.com/walkerke/neighborhood_diversity>

Political party colours: <http://blog.richardallen.co.uk/uk-political-party-web-colours/>

```{r, results = "asis"}
citationFn <- function(pkgName = "base") {
  cat(attr(unclass(citation(pkgName))[[1]], "textVersion"), "<br>", "<br>")
}
citationFn() 
citationFn("flexdashboard")
citationFn("shiny")
citationFn("tidyverse")
citationFn("readxl")
citationFn("scales")
citationFn("hrbrthemes")
citationFn("viridis")
citationFn("raster")
citationFn("leaflet")
```


<style>

.section.sidebar {
  background-color: white; 
}

 #muchSmaller { font-size: small; }

</style>
